A comparison of design-based and model-based approaches for finite population spatial sampling and inference.

### 2022-06-11 inst/manuscript/manuscript.Rmd
added zenodo reference
minor updates

The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships. First, the bounding box of the domain is split up into four distinct, equally sized squares called level-one cells. Each level-one cell is randomly assigned a level-one address of 0, 1, 2, or 3. The set of level-one cells is denoted by $\mathcal{A}_1$ and defined as $\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}$. Within each level-one cell, the inclusion probability for each population unit (which is pre-specified) is summed, and if any of these sums are one or greater, a second level of cells is added. Then each level-one cell is split into four distinct, equally sized squares called level-two cells. Each level-two cell is randomly assigned a level-two address of 0, 1, 2, or 3. The set of level-two cells is denoted by $\mathcal{A}_2$ and defined as $\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}$. The inclusion probabilities within each level-two cell are summed, and if any of these sums are one or greater, a third level of cells is added. This process continues for $k$ steps, until all level-$k$ cells have inclusion probability sums less than one. Then $\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}$. Figure$~$\ref{fig:grts} provides some intuition regarding the assignment of level-one and level-two cells.

@@

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation}
where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively, $\bm{\beta}$ is the parameter vector of fixed effects, and $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.

@@

\begin{figure}
\centering
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
  \caption{Histogram of a single realized population for the normal response.}
  \label{fig:symm_pop_hist}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
  \caption{Histogram of a single realized population for the skewed response.}
  \label{fig:skew_pop_hist}
\end{subfigure}
\caption{Histograms of single realized populations simulated for the normal and skewed respones using the random layout and 50\% DRE.}
\label{fig:sim_pops}
\end{figure}

@@

This manuscript has a supplementary \textbf{\textsf{R}} package that contains all of the data, code, and figures used in its creation. All figures were created using the ggplot2 \textbf{\textsf{R}} package [@wickham2016ggplot2]. The supplementary \textbf{\textsf{R}} package is hosted on GitHub and archived on Zenodo. The Zenodo DOI is located here and in the references.

>>>>

The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships. First, the bounding box of the domain is split up into four distinct, equally sized squares called level-one cells. Each level-one cell is randomly assigned a level-one address of 0, 1, 2, or 3. The set of level-one cells is denoted by $\mathcal{A}_1$ and defined as $\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}$. Within each level-one cell, the inclusion probability for each population unit (which is pre-specified) is summed, and if any of these sums are one or greater, a second level of cells is added. Then each level-one cell is split into four distinct, equally sized squares called level-two cells. Each level-two cell is randomly assigned a level-two address of 0, 1, 2, or 3. The set of level-two cells is denoted by $\mathcal{A}_2$ and defined as $\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}$. The inclusion probabilities within each level-two cell are summed, and if any of these sums are one or greater, a third level of cells is added. This process continues for $k$ steps, until all level-$k$ cells have inclusion probability sums less than one. Then $\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}$. Fig.$~$\ref{fig:grts} provides some intuition regarding the assignment of level-one and level-two cells.

@@

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation}
where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively, $\bm{\beta}$ is the parameter vector of fixed effects, and $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]^T$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.

@@

\begin{figure}
\centering
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
  \caption{Histogram of a single realized population for the normal response.}
  \label{fig:symm_pop_hist}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
  \caption{Histogram of a single realized population for the skewed response.}
  \label{fig:skew_pop_hist}
\end{subfigure}
\caption{Histograms of single realized populations simulated for the normal and skewed responses using the random layout and 50\% DRE.}
\label{fig:sim_pops}
\end{figure}

@@

This manuscript has a supplementary \textbf{\textsf{R}} package that contains all of the data, code, and figures used in its creation. All figures were created using the ggplot2 \textbf{\textsf{R}} package [@wickham2016ggplot2]. The supplementary \textbf{\textsf{R}} package is hosted on GitHub and archived on Zenodo. The Zenodo DOI link is https://doi.org/10.5281/zenodo.6633805 [@dumelle2022DvMsp].
